Network Configuration Synthesis with Abstract Topologies

### 2016-11-05 paper/abstraction/paper.tex
cleanup some fonts

Converting from the high-level syntax from \S\ref{sec:overview} to the core syntax is straightforward. For example, the predicate $\True$ becomes $0.0.0.0/[0..32]$. Preferences are lifted to the top level of the regular expression when unambiguous. For example, the constraint $\Exit(\sf{Peer1} \Prefer \sf{Peer2})$ becomes $\Exit(\sf{Peer1}) \Prefer \Exit(\sf{Peer2})$ before being rewritten into regular expressions. Each modular set of constraints are joined prefix-by-prefix by taking the intersection of their constraints. For example, the combining the Routing and Locality constraints in the data center policy from \S\ref{sec:overview} would result in the following \sysname policy:
%
\begin{code}
\$PG  \Path \End(TG),
\$PL  \Path \End(TL) \ensuremath{\cap} \Always(\In),
\True \Path \Exit(Peer1) \Prefer \Exit(Peer2)
\end{code}
\noindent%
%
Finally, each constraint desugars to a regular expression. For example, the constraint $\Always(\In)$ becomes $\In^*$ and the constraint $\End({\small \mathsf{T0}})$ becomes $\Sigma^* \cdot \sf{T0}$.

@@

For each regular expression $r_i$ from $r_1 ~\Prefer~ .. ~\Prefer~ r_k$, we construct a DFA for the reverse of $r_i$. An automaton for $r_i$ is defined as a tuple ($\Sigma, Q_i, F_i, q_{0_i}, \sigma_i$). The alphabet $\Sigma = G^A$ is the set of topology locations, $Q_i$ is the set of states for automaton i, $F_i$ is the set of final states, $q_{0_i}$ is the initial state, and $\sigma_i \colon Q_i \times \Sigma \rightarrow Q_i$ is the state transition function.
%
The Product Graph is a tuple ($G'$, $s$, $P$) where $G' = (V',E')$ is a new graph with
vertices $V' \colon V \times Q_1 \times \dots \times Q_j$,
edges $E' \colon V' \times V'$,
a unique starting vertex $s$,
and a preference function $P \colon V' \rightarrow 2^{\set{1, \dots, j}}$ , which maps nodes in the product graph to a set of path ranks.

@@

Figure~\ref{fig:example-compilation} shows an example of constructing a Product Graph for the policy that applies to all external traffic from the datacenter example in \S\ref{sec:overview}:
%
%\begin{code}
%\True \Path \Exit(\sf{Peer1}) \Prefer \Exit(\sf{Peer2})
%\end{code}
%\noindent%
%
The first automaton represents the more preferred constraint $\Exit(Peer1)$ and the second automaton represents the less preferred constraint $\Exit(Peer2)$. The Product Graph representation is shown for both an instance of a concrete network matching the abstraction from \S\ref{sec:overview} as well as for the abstraction itself.

@@

For example, in the concrete product graph we have $(S1,0,1) \leq (S1,1,0)$ since nodes on the left side of the product graph can always match transitions made on the right hand side while always maintaining a node ranking that is at least as good as that on the right hand side (since they match the preferred rank 1 automaton). This relation does not hold the other way around. Therefore, advertisements received at $S1$ in state $(1,0)$ must be preferred to those received in state $(0,1)$.

Notice also that in the abstract product graph, node $(S,0,1) \leq (S,1,0)$. This leads us the the following observation:

@@

Consider again the datacenter example from \S\ref{sec:overview}. For global prefixes we used the routing policy:
%
\[ \begin{array}{c}
  \$PG \Path end(TG)
\end{array} \]
\noindent
%
Suppose now we additionally add the valleyfree constraint that traffic never has takes a path with a down-up in it. Figure~\ref{fig:example-inference} shows the abstract product graph representation for this routing policy. The inference algorithm starts from the node $(TG,0,0)$ with the initial fact $S_P S_{TG}(1,\infty)$ (\IE, no restriction on the number of disjoint paths initially). 
%
The first step applies the rule \infrule{I-mesh1} together with \infrule{I-local} to reason about connectivity within a single pod for the TG and AG roles. This invokes a call to Z3Opt to minimize the expression $min(\infty, AG)$, which gives us 2. Therefore, the algorithm learns a new invariant of the form $S_P A_{AG} (2,1)$ for node $(AG, 0)$ to indicate that in some pod P, any group of 2 nodes is reachable.
%
The next step uses the inference rule \infrule{I-out2} to learn that any single spine node is reachable. It also can apply rule \infrule{I-striping} to determine that there is some group of at least 2 spine routers reachable at node $(S,0)$ and also that there is some group of at least 2 nodes reachable in the AL role in state $(AL,0)$.

\ryan{inference should only infer some}

>>>>

Converting from the high-level syntax from \S\ref{sec:overview} to the core syntax is straightforward. For example, the predicate $\True$ becomes $0.0.0.0/[0..32]$. Preferences are lifted to the top level of the regular expression when their use is unambiguous. For example, the constraint $\Exit\CD{(Peer1} ~\Prefer~ \CD{Peer2)}$ becomes $\Exit\CD{(Peer1)} \Prefer~ \Exit\CD{(Peer2)}$ before being rewritten into regular expressions. Each modular set of constraints are joined prefix-by-prefix by taking the intersection of their constraints. For example, the combining the Routing and Locality constraints in the data center policy from \S\ref{sec:overview} would result in the following \sysname policy:
%
\begin{code}
\$PG  \Path \End(TG),
\$PL  \Path \End(TL) \ensuremath{\cap} \Always(\In),
\True \Path \Exit(Peer1) \Prefer \Exit(Peer2)
\end{code}
\noindent%
%
Finally, each constraint desugars to a regular expression. For example, the constraint $\Always\CD{(\In)}$ becomes $\In^*$ and the constraint $\CD{\End(T0)}$ becomes $\mathtt{\Sigma}^* \cdot \CD{T0}$.

@@

For each regular expression constraint $r_i$ from $r_1 ~\Prefer~ .. ~\Prefer~ r_k$, we construct a DFA for the reverse of $r_i$. An automaton for $r_i$ is defined as a tuple ($\Sigma, Q_i, F_i, q_{0_i}, \sigma_i$). The alphabet $\Sigma = G^A$ is the set of topology locations for the abstraction, $Q_i$ is the set of states for automaton i, $F_i$ is the set of final states, $q_{0_i}$ is the initial state, and $\sigma_i \colon Q_i \times \Sigma \rightarrow Q_i$ is the state transition function.
%
The Product Graph is a tuple ($G'$, $s$, $P$) where $G' = (V',E')$ is a new graph with
vertices $V' \colon V \times Q_1 \times \dots \times Q_j$,
edges $E' \colon V' \times V'$,
a unique starting vertex $s$,
and a preference function $P \colon V' \rightarrow 2^{\set{1, \dots, j}}$ , which maps nodes in the product graph to a set of path ranks.

@@

Figure~\ref{fig:example-compilation} shows an example of constructing a Product Graph for the policy that applies to all external traffic from the datacenter example in \S\ref{sec:overview}:
%
%\begin{code}
%\True \Path \Exit(\sf{Peer1}) \Prefer \Exit(\sf{Peer2})
%\end{code}
%\noindent%
%
The first automaton represents the more preferred constraint $\CD{\Exit(Peer1)}$ and the second automaton represents the less preferred constraint $\CD{\Exit(Peer2)}$. The Product Graph representation is shown for both an instance of a concrete network matching the abstraction from \S\ref{sec:overview} as well as for the abstraction itself.

@@

For example, in the concrete product graph the inequality $(S1,0,1) \leq (S1,1,0)$ holds since nodes on the left side of the product graph can always match transitions made on the right hand side while always maintaining a node ranking that is at least as good as that on the right hand side (since they match the preferred rank 1 automaton). This relation does not hold the other way around. Therefore, advertisements received at $S1$ in state $(1,0)$ must be preferred to those received in state $(0,1)$.

Notice also that in the abstract product graph the inequality $(S,0,1) \leq (S,1,0)$ holds. This leads us the the following observation:

@@

Consider again the datacenter example from \S\ref{sec:overview}. Suppose now we add the valleyfree constraint that traffic never has takes a path with a down-up in it for the routing policy for global prefixes:
%
\begin{code}
\$GP  \Path \End(TG) \ensuremath{\cap} \Novalley(\{TG,TL\},\{AG,AL\},\{S\})
\end{code}
\noindent
%
Figure~\ref{fig:example-inference} shows the abstract product graph representation for this routing policy. The inference algorithm starts from the node $(TG,0,0)$ with the initial fact $S_P S_{TG}(1,\infty)$ (\IE, no restriction on the number of disjoint paths initially). 
%
The first step applies the rule \infrule{I-mesh1} together with \infrule{I-local} to reason about connectivity within a single pod for the TG and AG roles. This invokes a call to Z3Opt to minimize the expression $min(\infty, AG)$, which gives us 2. Therefore, the algorithm learns a new invariant of the form $S_P A_{AG} (2,1)$ for node $(AG, 0)$ to indicate that in some pod P, any group of 2 nodes is reachable.
%
The next step uses the inference rule \infrule{I-out2} to learn that any single spine node is reachable. It also can apply rule \infrule{I-striping} to determine that there is some group of at least 2 spine routers reachable at node $(S,0)$ and also that there is some group of at least 2 nodes reachable in the AL role in state $(AL,0)$.
